### 2007-01-14 Documentation/part1.tex
JK - Minor changes in chapters 8 - 9

\GLOBES\ 3.0 and higher does not require a re-compilation of the software to simulate non-standard physics.
However, the probability engine has to be changed, which one would usually do by copying the respective parts from {\tt glb\_probability.c} (in the {\tt source} sub-directory). Therefore, for sophisticated applications, familiarity with this file is required. A number of comments are provided in this file to keep it self-explanatory.
The key function types are
\begin{function}
\GLBNS{\_set\_oscillation\_parameters\_function}
{\tt int (*glb\_set\_oscillation\_parameters\_function)(glb\_params p, void* user\_data)} reads the fundamental oscillation parameters
obtained from {\tt p} and stores them into global variables accessible to {\tt glb\_probability\_matrix\_function}.
 In addition, it can be used to pre-compute quantities which are dependent on the oscillation paramaters only for efficiency. 
In order to circumvent global variables, an arbitrary pointer can be used by {\tt user\_data} which is set by {\tt glbRegisterProbabilityEngine}.
\end{function}
\begin{function}
\GLBNS{\_get\_oscillation\_parameters\_function}
{\tt int (*glb\_get\_oscillation\_parameters\_function)(glb\_params p, void* user\_data)} returns the fundamental oscillation parameters
in {\tt p} which are obtained from the global variables accessible to {\tt glb\_probability\_matrix\_function}.
In order to circumvent global variables, an arbitrary pointer can be used by {\tt user\_data} which is set by {\tt glbRegisterProbabilityEngine}.
\end{function}
\begin{function}
\GLBNS{\_probability\_matrix\_function}
{\tt int (*glb\_probability\_matrix\_function)(double P[3][3], int cp\_sign, double E,
                  int psteps, const double *lengths, const double *densities,
                  double filter\_sigma, void* user\_data)} calculates the neutrino oscillation probability matrix to be returned in {\tt P}. The {\tt cp\_sign} is +1 for neutrinos and -1 for antineutrinos. In addition, the matter density profile to be used is characterized by the number of steps {\tt psteps}, the lengths of the matter density layers in the list {\tt lengths}, and the corresponding densities in the list {\tt densities}. The parameter {\tt filter\_sigma} defines the width of the low-pass filter to be used, or no filter usage if negative.
In order to circumvent global variables, an arbitrary pointer can be used by {\tt user\_data} which is set by {\tt glbRegisterProbabilityEngine}.
\end{function}
These three function types correspond to the standard implementation functions \GLB{\_set\_oscillation\_parameters}, \GLB{\_get\_oscillation\_parameters}, and \GLB{\_probability\_matrix} in {\tt glb\_probability.c} of the source code of \GLOBES , where one can find the standard behavior and use it by cut and paste (including the variables and code used by these functions).

In order to use non-standard physics in \GLOBES , one needs re-define the three above functions and register them after {\tt glbInit} with
\begin{function}
\GLBNS{RegisterProbabilityEngine}
{\tt int glbRegisterProbabilityEngine(int n\_parameters,
                 glb\_probability\_matrix\_function prob\_func, \\
                 glb\_set\_oscillation\_parameters\_function set\_params\_func, \\
                 glb\_get\_oscillation\_parameters\_function get\_params\_func, \\
                 void* user\_data)} registers a probability
engine for the simulation of non-standard physics with {\tt n\_parameters} oscillation parameters ({\tt n\_parameters} $\ge6$). The three functions {\tt prob\_func}, {\tt set\_params\_func}, and {get\_params\_func} 
will be used with the function types defined above. In order to circumvent global variables, an arbitrary pointer can be used by {\tt user\_data} being set by this function.
\end{function}
The number of oscillation parameters can, at any time, be obtained with
\begin{function}
\GLBNS{GetNumOfOscParams}
{\tt  int glbGetNumOfOscParams()} returns the number of oscillation parameters.
\end{function}

@@

or any other such effect, \ie, one more oscillation parameter. We need to define two functions to access a set of 

@@

\example{Decoherence in neutrino oscillations}{
\label{ex:nsphysics}
\index{norm}{Decoherence}
The following code fragment from {\tt example6.c} calculates 
a fit region on $\stheta$-$\sigma_E$-space, where $\sigma_E$ is the
non-standard decoherence parameter (\cf, \Ref~\cite{Blennow:2005yk}):
\begin{quote}
{\tt {\footnotesize
  const GLB\_SIGMA\_E = 6; \\ \\
  /* Fix theta13 and sigma\_E, and marginalize over all other parameters except deltaCP and th23, which do not enter P\_ee */ \\
  glbDefineProjection(myproj, GLB\_FREE, GLB\_FIXED, GLB\_FIXED,
    GLB\_FIXED, \\ \hspace*{0.5cm} GLB\_FREE, GLB\_FREE); \\
  glbSetDensityProjectionFlag(myproj, GLB\_FIXED, GLB\_ALL); \\
  glbSetProjectionFlag(myproj, GLB\_FIXED, GLB\_SIGMA\_E);  \\
  glbSetProjection(myproj); \\ 
\\
  for(x=0; x < 0.05+0.001; x+=0.005)      /* th13 loop */ \\
  for(y=0.0; y < 0.010+0.001; y+=0.001)   /* sigma\_E loop */ \\
    \{ \\
   \hspace*{0.5cm}    /* Set vector of test=fit values */ \\
   \hspace*{0.5cm}    thetheta13=asin(sqrt(x))/2.0; \\
   \hspace*{0.5cm}    glbSetOscParams(test\_values,thetheta13,GLB\_THETA\_13); \\
   \hspace*{0.5cm}    glbSetOscParams(test\_values,y,GLB\_SIGMA\_E); \\
     \\
   \hspace*{0.5cm}    /* Compute Chi2 with correlations */ \\
   \hspace*{0.5cm}    res=glbChiNP(test\_values,NULL,GLB\_ALL); \\
  \\
   \hspace*{0.5cm}    AddToOutput(x,y,res); \\
    \} 
}}
\end{quote}

@@

Here we describe experimental features currently being implemented in \GLOBES . These features have not yet been tested extensively and should be used with care. They may evolve into standard features in the future, or they may not be supported anymore at a certain point. In general, the number of experiment features is small in a new release version, and increases towards a new version number.

>>>>

\GLOBES\ 3.0 and higher does not require a re-compilation of the software to simulate non-standard physics.
However, the probability engine has to be changed, which one would usually do by copying the respective parts from {\tt glb\_probability.c} (in the {\tt source} sub-directory). Therefore, for sophisticated applications, familiarity with this file is required. A number of comments are provided in this file to keep it self-explanatory.
The key function types are
\begin{function}
\GLBNS{\_set\_oscillation\_parameters\_function}
{\tt int (*glb\_set\_oscillation\_parameters\_function)(glb\_params p, void* user\_data)} is
used to pass the fundamental oscillation parameters {\tt p} to the probability engine.
The function should store them into variables accessible to {\tt glb\_probability\_matrix\_function}.
In addition, it can be used to pre-compute quantities which are dependent on the oscillation paramaters for efficiency. 
In order to circumvent global variables, arbitrary additional parameters can be
passed to the function in {\tt user\_data}, which is set by {\tt glbRegisterProbabilityEngine}.
\end{function}
\begin{function}
\GLBNS{\_get\_oscillation\_parameters\_function}
{\tt int (*glb\_get\_oscillation\_parameters\_function)(glb\_params p, void* user\_data)}
reads the fundamental oscillation parameters from the internal variables of the probability
engine, and writes them into {\tt p}.
In order to circumvent global variables, arbitrary additional parameters can be
passed to the function in {\tt user\_data}, which is set by {\tt glbRegisterProbabilityEngine}.
\end{function}
\begin{function}
\GLBNS{\_probability\_matrix\_function}
{\tt int (*glb\_probability\_matrix\_function)(double P[3][3], int cp\_sign, double E,
                  int psteps, const double *lengths, const double *densities,
                  double filter\_sigma, void* user\_data)}
calculates the neutrino oscillation probability matrix and returns it in {\tt P}.
The {\tt cp\_sign} is +1 for neutrinos and -1 for antineutrinos. In addition, the matter
density profile to be used is characterized by the number of steps {\tt psteps}, the
lengths of the matter density layers in the list {\tt lengths}, and the corresponding
densities in the list {\tt densities}. The parameter {\tt filter\_sigma} defines
the width of the low-pass filter to be used, or no filter usage if negative.
In order to circumvent global variables, arbitrary additional parameters can be
passed to the function in {\tt user\_data}, which is set by {\tt glbRegisterProbabilityEngine}.
\end{function}
These three function types correspond to the standard implementation functions \GLB{\_set\_oscillation\_parameters}, \GLB{\_get\_oscillation\_parameters}, and \GLB{\_probability\_matrix} in {\tt glb\_probability.c} of the source code of \GLOBES , where one can find the standard behavior and use it by cut and paste (including the variables and code used by these functions).

In order to use non-standard physics in \GLOBES , one needs to re-define the three above functions and register them after {\tt glbInit} with
\begin{function}
\GLBNS{RegisterProbabilityEngine}
{\tt int glbRegisterProbabilityEngine(int n\_parameters,
                 glb\_probability\_matrix\_function prob\_func, \\
                 glb\_set\_oscillation\_parameters\_function set\_params\_func, \\
                 glb\_get\_oscillation\_parameters\_function get\_params\_func, \\
                 void* user\_data)} registers a probability
engine for the simulation of non-standard physics with {\tt n\_parameters} oscillation parameters ({\tt n\_parameters} $\ge6$). The three functions {\tt prob\_func}, {\tt set\_params\_func}, and {\tt get\_params\_func} 
will be used with the function types defined above. In order to circumvent global variables, an arbitrary pointer {\tt user\_data} can be defined, that will be passed to the probability engine in each function call.
\end{function}
The number of oscillation parameters can, at any time, be obtained with
\begin{function}
\GLBNS{GetNumOfOscParams}
{\tt  int glbGetNumOfOscParams()} returns the number of oscillation parameters.
\end{function}

@@

or any other such effect, \ie, we have one additional oscillation parameter. We need to define two functions to access a set of 

@@

\example{Decoherence in neutrino oscillations}{
\label{ex:nsphysics}
\index{norm}{Decoherence}
The following code fragment from {\tt example6.c} calculates 
a fit region on $\stheta$-$\sigma_E$-space, where $\sigma_E$ is the
non-standard decoherence parameter (\cf, \Ref~\cite{Blennow:2005yk}):
\begin{quote}
{\tt {\footnotesize
  const int GLB\_SIGMA\_E = 6; \\ \\
  /* Fix theta13 and sigma\_E, and marginalize over all other parameters except deltaCP and th23, which do not enter P\_ee */ \\
  glbDefineProjection(myproj, GLB\_FREE, GLB\_FIXED, GLB\_FIXED,
    GLB\_FIXED, \\ \hspace*{0.5cm} GLB\_FREE, GLB\_FREE); \\
  glbSetDensityProjectionFlag(myproj, GLB\_FIXED, GLB\_ALL); \\
  glbSetProjectionFlag(myproj, GLB\_FIXED, GLB\_SIGMA\_E);  \\
  glbSetProjection(myproj); \\ 
\\
  for(x=0; x < 0.05+0.001; x+=0.005)      /* th13 loop */ \\
  for(y=0.0; y < 0.010+0.001; y+=0.001)   /* sigma\_E loop */ \\
    \{ \\
   \hspace*{0.5cm}    /* Set vector of test=fit values */ \\
   \hspace*{0.5cm}    thetheta13=asin(sqrt(x))/2.0; \\
   \hspace*{0.5cm}    glbSetOscParams(test\_values,thetheta13,GLB\_THETA\_13); \\
   \hspace*{0.5cm}    glbSetOscParams(test\_values,y,GLB\_SIGMA\_E); \\
     \\
   \hspace*{0.5cm}    /* Compute Chi2 with correlations */ \\
   \hspace*{0.5cm}    res=glbChiNP(test\_values,NULL,GLB\_ALL); \\
  \\
   \hspace*{0.5cm}    AddToOutput(x,y,res); \\
    \} 
}}
\end{quote}

@@

Here we describe experimental features currently being implemented in \GLOBES . These features have not yet been tested extensively and should be used with care. They may evolve into standard features in the future, or they may not be supported anymore at a certain point. In general, the number of experimental features is small in a new release version, and increases towards a new version number.
